---
title: "Regression Analysis Salary Dataset"
author: "Gilbert Tarus"
date: "2023-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

## Loading the Packages

```{r}
library(dplyr)
library(caTools)
library(ggplot2)
library(readr)
library(skimr)
library(naniar)
library(caret)
```

## Reading Data

```{r}
salary <- read_csv("data/Salary_dataset.csv")
```
## Data Cleaning

```{r}
knitr::kable(head(salary))
```


```{r}
salary %>%  select(-1) -> salary
```

## Exploratory Analysis

```{r}
str(salary)
```
```{r}
salary %>% skim()
```
### Data Visualization

```{r}
## Scatter plot
ggplot(salary, aes(x = YearsExperience, y = Salary))+
  geom_point()+
  geom_smooth(se = FALSE, method = "loess")
```

### Summary Statistics

```{r}
salary %>% summary()
```
### Correlation Analysis

```{r}
salary %>% 
  cor()
```
There is a very string positive relationship between the Years of Experience and Salary.

## Data Selection

```{r}
## Training and test data
set.seed(2023-08-21-16)
sample_index <- caTools::sample.split(salary$Salary, SplitRatio = .65)
train <- subset(salary, sample_index)
test <- subset(salary, !sample_index)

nrow(train)
nrow(test)
```
## Model

```{r}
mod1 <- lm(Salary ~., data = train)
summary(mod1)
```
From the results above, Years of Experience have a very small p-value. This makes it more significant to predict salary.

Adjusted R-Squared is ```r summary(mod1)[["adj.r.squared"]]```. This indicates that the model  explains more that ```r summary(mod1)[["adj.r.squared"]]*100```% of the variability in the data. This is indicated by a very small p-value.

## Prediction

```{r}
predictions <- predict(mod1, newdata = test)
actual_prediction <- cbind(Salary = test$Salary, Predicted = predictions)
actual_prediction <- as.data.frame(actual_prediction)
actual_prediction <- actual_prediction %>% mutate(error = Salary - Predicted)
summary(actual_prediction$error)
```
Conclusively, the model performance is ```r summary(mod1)[["adj.r.squared"]]*100```%. This implies that the model can explain ```r summary(mod1)[["adj.r.squared"]]*100```% of the variability in the data.


## The model

Salary = ```r options(scipen = 999); round(summary(mod1)[["coefficients"]][1],3)``` + ```r round(summary(mod1)[["coefficients"]][2],2)``` * Years of Experience

**Question**

1. Predict the salary of an employee with 10 years of experience

```{r}
predict(mod1, data.frame(YearsExperience = 10))
```

